Learning by Teaching a Robot:\\The Case of Handwriting

### 2015-07-10 main.tex
Reworked Diego's study part

We present in this article a new take on this educative challenge, as a
remediation procedure that involve a ``bad writer'' robot that is \emph{taught}
by the child: by building on the \emph{learning by teaching} paradigm, not only
the child practises hand-writing but, as (s)he takes on the role of the teacher,
(s)he also positively reinforce his/her self-esteem and motivation: his/her
social role shifts from the ``underperformer'' to ``the one who knows and
teaches''.

@@

Further qualitative assessment of engagement over longer interactions were
conducted during the case studies, as reported in the following section.

@@

\subsubsection{Context, Study Design}
The CoWriter activity needs a child engaged as interaction leader. In this
study we consider the problem of long-term interactions : is it possible to sustain this
engagement over several one-hour sessions ?\\

Vincent is a six years old child with troubles in handwriting and a poor
self-esteem. Our goal was to figure out an environment in
order to make Vincent sustaining engagement over four sessions of one hour, one
session per week. We decided to introduce an appealing scenario that justifyed to the child
the activity where a robot wants to learn handwriting. We used two Nao robots :
a blue one (called Mimi) and an orange one (called Clem). Mimi was away for a
mission,and the two robots had to communicate by mails. But they decided to do
it ``like humans", with handwritten messages. While Mimi was a good in
handwriting, Clem had strong difficulties and needed the help of Vincent.\\

During the first session, Vincent just teached single letters to the robot.
To chose a letter, he had to pick it in a set of magneted plastic letters, and to
show it to the robot. But the robot was not programmed to detect and recognize
it. A third person was entering the letter and was sending it to the
system with the computer. With following sessions we gradually increased the
difficulty : in the second session the child teached full words. To chose the
word, he had to assemble plastic letters to build it and to show it to the
robot. In the two last sessions, he teached long sentences. This time he just
had to say the sentence to the robot, and then the third person was using the
compurter to enter it. \\

We designed a special scenario for the last session : Mimi came back to the
laboratory. In front of Vincent, Mimi said to Clem that it did not believe 
it realy improved its handwritting, and that the responses it recieved
was written by Vincent. To prove that it actualy improved its writing,
it gave to Clem thirty minutes for a final training. After the training, it
asked to Clem to write some words. Finally, he admitted that Clem actually
improved its handwriting.

\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{diego}
    \caption{\small Vincent working with Nao}
    \label{fig:diego}
\end{figure}

\subsubsection{Results}
In all sessions, Vincent was focused on the activity during forty minutes up to
one hour. In total, he provided 154 letters as demonstration.\\

Vincent accepted our scenario and seriously played the game. At the
beginning he seemed not confident and intimidated.  But quickly, he gain in
confidence to play with the robot. He even showed real affective bonds : he
sent an handwritten mail to it some days after the last session.\\

As a technical result, we do not know if Vincent improved his handwriting but he
likely conviced himself he was good enough at writing to help someone else (the
robot in that case). To show to Vincent how much he improved the robot's
handwriting, we printed the letter that the robot wrote during the third session
before and after his help (Fig.~\ref{fig:stimuli}).

@@

    \caption{\small Text generated by the robot, before and after training with
        the child (in French). As an example, the red box emphasizes the changes on the word
    ``envoyer''.}
    \label{fig:stimuli}
\end{figure}

@@

    \item this proves essential to \textbf{sustain a long-term interaction} (several
        hours) around a fundamentally routine -- yet challenging -- educational
        task (hand-writing learning),

>>>>

We present in this article a new take on this educative challenge, as a
remediation procedure that involve a ``bad writer'' robot that is \emph{taught}
by the child: by building on the \emph{learning by teaching} paradigm, not only
the child practises handwriting but, as (s)he takes on the role of the teacher,
(s)he also positively reinforce his/her self-esteem and motivation: his/her
social role shifts from the ``underperformer'' to ``the one who knows and
teaches''.

@@

We also performed further qualitative assessment of engagement over longer
interaction periods: two 4-weeks long case studies with one child at a time have
been conducted so far, the first one (\emph{Vincent}) focusing on
\emph{engagement on the long run}, the second one (\emph{Thomas}) looking at
\emph{acceptance and impact in a real therapeutic context}.

@@

During this first case study, we invited and followed Vincent, a six years old
child, once per week over a period of a month. Our primary aim was to address the
question of whether we can sustain Vincent's engagement and commitment to the
writing activity over such a period.

\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{diego}
    \caption{\small Vincent correcting {\sc nao}'s attempt by rewriting the
        whole word. Empty boxes are drawn on the screen to serve as template for the child
        and to make word segmentation more robust.}
    \label{fig:diego}
\end{figure}

The study took place at our laboratory (Figure~\ref{fig:diego}), and we chose to
design the activity around a storyline meant to be attractive for a 6 years old
boy: one of our {\sc nao} was away for a mysterious scientific mission, and it
needed the support of another one -- which stayed at the lab -- to interpret
curious pictures that were sent every week. Vincent had to help the second robot
understanding the pictures, and since the two robots had somehow beforehand
agreed to communicate with ``letters written like humans'' (\ie handwritten),
Vincent also had to help the robot to write good looking letters (because, well,
this robot was terrible at writing!) The experimental setup was similar to
Figure~\ref{experimental_setup}, except that Vincent had to tell the robot what
to write with small plastic letters (visible behind the robot on Figure~\ref{fig:diego}).

To complement the intrinsic motivation of helping a robot to communicate with another one, we
gradually increased the complexity of Vincent's task to keep it challenging and
interesting (the first week: demonstration of single letters; the second week:
short words; the third week: a full letter -- Figure~\ref{fig:stimuli}).

The last session was set as a test: the ``explorer'' robot
had come back from its mission and it actually challenged the other robot in
front of Vincent: \emph{``I don't believe you wrote yourself these nice letters that I
received! Prove it to me by writing something in front of me!''} This situation
was meant to evidence the Prot\'eg\'e effect: by judging the other robot's
handwriting, the ``explorer'' robot would implicitly judge Vincent's skills as
teacher, and in turn, Vincent's handwriting.

@@

    \caption{\small (French) text generated by the robot, before and after a one
        hour long interaction session with the child. As an example, the red box
        highlights the changes on the word ``envoyer''.}

    \label{fig:stimuli}
\end{figure}

Overall, Vincent provided 154 demonstrations to the robot, and he remained
actively engaged over the four weeks. The story was well accepted by Vincent and
he seriously engaged into the game. After the first week, he showed good
confidence to play with the robot and he built affective bonds with the robot
over the course of the study, as evidenced by some cries on the last session,
and several letters sent by him to the robot \emph{after} the end of the study
(one of them 4 months later) to get news. This represents a promising initial
result: we can effectively keep a child engaged with the robot for a relatively
long period of time (about 5 hours).

No conclusion can be drawn in terms of actual handwriting remediation: we did
not design this study to formally assess possible improvements.

However, as pictured on Figure~\ref{fig:stimuli}, Vincent was able to
significantly improve the robot's skill, and he acknowledged that he had been
able to help the robot: in that regard, Vincent convinced himself that he was
``good enough'' at writing to help someone else, and this is likely to have
positively impacted his self-esteem.

@@

    \item this proves essential to \textbf{sustain a long-term interaction} (several
        hours) around a fundamentally routine -- yet challenging -- educational
        task (handwriting learning),
